---
title: "Project1_TimeSeriesAnalysis"
author: "Matt Chinchilla"
date: "2/11/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tseries)
library(forecast)
library(car)
library(TSstudio)
```


```{r}
#setwd("/Users/matt/Desktop/SMU/Applied Stats/ProjectDetails/AppliedStats")
BitcoinDataMergeV2 <- read.csv("BitcoinDataMergeV2.csv", header = TRUE)
```

Add a date column per Project requirements
```{r}
colnames(BitcoinDataMergeV2)[1] <- "Time"
```

Plot Bitcoin Price data over time
```{r, plot}
p <- ggplot(BitcoinDataMergeV2, aes(x=Time, y=Price)) +
  geom_line() + 
  xlab("Time") + ylab("Bitcoin Price") + ggtitle("Bitcoin Price Over Time")
p
```

```{r}
attach(BitcoinDataMergeV2)
acf(Price)
```

```{r}
pacf(Price)
```
Run durbin watson determin if serial corrolation exists 
```{r}
durbinWatsonTest(lm(Price~1), max.lag = 10)
```

Convert to time series
```{r}
bitcoinTS_Price <- ts(BitcoinDataMergeV2$Price, freq=365)
```


Decompose Time Series
```{r}
bitcoin_tsDecomp <- decompose(bitcoinTS_Price, type = 'multiplicative')
plot(bitcoin_tsDecomp)
```


See what data has correlation 
```{r}
library(corrplot)
library(purrr) #keep
        
#function to create corrolation heatmap
correlator <- function(df){
  df %>%
    keep(is.numeric) %>%
    tidyr::drop_na() %>%
    cor %>%
    corrplot(addCoef.col = "white", number.digits = 2,
             number.cex = .5, method = "pie",
             order = "hclust",
             tl.srt = 45, tl.cex = .8)
}

correlator(BitcoinDataMergeV2)
```

Log Transform price data to compare
```{r}
BitcoinDataMergeV2$LogPrice <- log(Price)
attach(BitcoinDataMergeV2)
```


Plot Log transformed Bitcoin Price data over time
```{r, plot}
p <- ggplot(BitcoinDataMergeV2, aes(x=Time, y=LogPrice)) +
  geom_line() + 
  xlab("Time") + ylab("Bitcoin Price") + ggtitle("Log transformed Bitcoin Price Over Time")
p
```

```{r}
acf(LogPrice)
```

```{r}
pacf(LogPrice)
```


```{r}
bitcoinTS_LogPrice <- ts(LogPrice, freq=365)
```

Decompose Time Series
```{r}
bitcoin_tsDecomp2 <- decompose(bitcoinTS_LogPrice, type = 'multiplicative')
plot(bitcoin_tsDecomp2)
```

Analysis of "Total BTC" as a Price Predictor
```{r}
plot(`Total BTC`,LogPrice,xlab="Total BTC")
Pred_TotalBTC<-lm(LogPrice~ poly(`Total BTC`,degree = 3,raw = TRUE))
Pred_TotalBTC2<-lm(LogPrice~ poly(`Total BTC`,degree = 2,raw = TRUE))
abline(lm(LogPrice~`Total BTC`))
lines(smooth.spline(`Total BTC`,predict(Pred_TotalBTC)), col = 'blue', lwd = 3)
text(16000000,5,paste("Cor=",round(cor(LogPrice,`Total BTC`),2)))
```
Polynomial model ^3
```{r}
summary(Pred_TotalBTC)
```

Polynomial model ^2
```{r}
summary(Pred_TotalBTC2)
```

Analysis of "Difficulty" as a Price Predictor
```{r}
plot(Difficulty,LogPrice,xlab="Total BTC")
Pred_Difficulty<-lm(LogPrice~ poly(Difficulty,degree = 2,raw = TRUE))
abline(lm(LogPrice~Difficulty))
lines(smooth.spline(Difficulty,predict(Pred_Difficulty)), col = 'blue', lwd = 3)
text(800000000000,5,paste("Cor=",round(cor(LogPrice,Difficulty),2)))
```

summary of Difficulty simple linear model
```{r}
DifficultyLM <- lm(LogPrice~Difficulty)
summary(DifficultyLM)
```

Summary of Difficulty polynomial fit ^2
```{r}
summary(Pred_Difficulty)
```

Compare Difficulty simple model to polynomial model: P <.05 there is a significant difference between these models  we'll include the ploynomial set in the predictors
```{r}
anova(DifficultyLM,Pred_Difficulty)
```


Analysis of "Count" as a Price Predictor
Conculsion: Count and Total BTC are not statistically different but count gets a slightly better r2 so we will use count as a predictor in our model instead of Total BTC
```{r}
plot(Count,LogPrice,xlab="Total BTC")
Pred_Count2<-lm(LogPrice~ poly(Count,degree = 2,raw = TRUE))
Pred_Count3<-lm(LogPrice~ poly(Count,degree = 3,raw = TRUE))
abline(lm(LogPrice~Count))
lines(smooth.spline(Count,predict(Pred_Count3)), col = 'blue', lwd = 3)
text(450000,5,paste("Cor=",round(cor(LogPrice,Count),2)))
```



```{r}
summary(Pred_Count3)
```

Analysis of "MarketCap" as a Price Predictor
```{r}
plot(Market.Cap,LogPrice,xlab="Total BTC")
Pred_MC1<-lm(LogPrice~Market.Cap)
Pred_MC2<-lm(LogPrice~ poly(Market.Cap,degree = 2,raw = TRUE))
Pred_MC3<-lm(LogPrice~ poly(Market.Cap,degree = 3,raw = TRUE))

abline(lm(LogPrice~Market.Cap))
lines(smooth.spline(Market.Cap,predict(Pred_MC3)), col = 'blue', lwd = 3)
text(800000000000,5,paste("Cor=",round(cor(LogPrice,Market.Cap),2)))
```

```{r}
summary(Pred_MC3)
```

Create a matrix of predictors to use
```{r}
train_preds <- as.matrix(cbind(Count[1:1467],Count[1:1467]^2,Count[1:1467]^3,Difficulty,Difficulty[1:1467]^2))
colnames(Preds)[2] <- c("Count^2")
colnames(Preds)[3] <- c("Count^3")
colnames(Preds)[5] <- c("Difficulty^2")


```

Split Time series
```{r}
Split_Data <- ts_split(ts.obj = bitcoinTS_LogPrice, sample.out = 730)
train <- Split_Data$train
test <- Split_Data$test

length(train)
length(test)
```

```{r, like hw4}
holdout.test<-window(ts(LogPrice),start=1468)
train<-window(ts(LogPrice, end = 1467, frequency = 90))
predictor1<-Count[1:1467]
predictor2 <- Difficulty[1:1467]
predictor3 <- Market.Cap[1:1467]

Preds <- as.matrix(cbind(predictor1,predictor1^2,as.numeric(as.character(predictor1^3)),predictor2,predictor2^2))
colnames(Preds)<-c("Count","Count2","Count3","Difficulty","Difficulty2")

Preds <- read.csv("Preds.csv")
Preds <- as.matrix(Preds)
```

Split Predictors
```{r}
train_preds <- slice(Preds2,1:1467)
test_preds <- slice(Preds2,1468:2197)

#train_prds <- as.matrix(train_preds)

length(train_preds)
length(test_preds)
```


#Model 1 - no predictors
```{r}
BTC_Model1<-auto.arima(train,stepwise=FALSE)
BTC_Model1
tsdisplay(residuals(BTC_Model1),lag.max=15,main="BTC_Model1")
plot(forecast(BTC_Model1,h=180))
points(1:length(train),fitted(BTC_Model1),type="l",col="blue")
```

#Model 2 - count predictor
```{r}
BTC_Model2<-auto.arima(train,xreg = predictor1, stepwise=FALSE)
BTC_Model2

tsdisplay(residuals(BTC_Model2),lag.max=15,main="BTC_Model2")

plot(forecast(BTC_Model2,h=90, xreg= as.matrix(data.frame(predictor1 = Count[1468:2197])))) 
#plot(forecast(ARIMA.with.Pred,h=5,xreg= as.matrix(data.frame(predictor = AvgTemp[36:40]))))
points(1:length(train),fitted(BTC_Model2),type="l",col="blue")
```

#Model 3 difficulty predictor
```{r}
BTC_Model3<-auto.arima(train,xreg = predictor2, stepwise=FALSE)
BTC_Model3
tsdisplay(residuals(BTC_Model3),lag.max=15,main="BTC_Model1")
plot(forecast(BTC_Model3, xreg = predictor2,h=90))
points(1:length(train),fitted(BTC_Model3),type="l",col="blue")
```

#Model 4 market cap
```{r}
BTC_Model4<-auto.arima(train,xreg = predictor3, stepwise=FALSE)
BTC_Model4
tsdisplay(residuals(BTC_Model4),lag.max=15,main="BTC_Model1")
plot(forecast(BTC_Model4,h=180))
points(1:length(train),fitted(BTC_Model4),type="l",col="blue")
```

#AR models
```{r}
Bitcoin_AR1 <- arima(LogPrice, order=c(1,0,0))

tsdisplay(residuals(Bitcoin_AR1),lag.max=5,main="AR(1) Resid. Diagnostics")
```


```{r}
Bitcoin_AR2 <- arima(LogPrice, order=c(2,0,0))
Bitcoin_AR2
tsdisplay(residuals(Bitcoin_AR2),lag.max=15,main="AR(2) Resid. Diagnostics")
```

```{r}
Bitcoin_AR3 <- arima(LogPrice, order=c(3,1,2))
Bitcoin_AR3
tsdisplay(residuals(Bitcoin_AR3),lag.max=15,main="AR(3) Resid. Diagnostics")
```


```{r}
Bitcoin_AutoFit <- auto.arima(BitcoinDataMergeV2, stepwise = FALSE)
Bitcoin_AutoFit
```

```{r}
tsdisplay(residuals(Bit),lag.max=15,main="AR(2) Resid. Diagnostics")
```

```{r}
plot(forecast(Bitcoin_AutoFit,h=90))
points(1:length(bitcoin_ts),fitted(Bitcoin_AutoFit),type="l",col="blue")
```